Alone Together: Compositional Reasoning and Inference for Weak Isolation

### 2017-11-06 inference.tex

The state transformer rules, like the earlier RG rules, closely follow
the corresponding reduction rules in Fig.~\ref{fig:txnimp}, except
that their language of expression is $\SL$. For instance, while the
reduction rule for \C{UPDATE} declaratively specifies the set of
updated records, the state transformer rule uses $\SL$'s bind
operation to \emph{compute} the set. Other SQL rules do likewise. The
rules for \C{LET} binders, conditionals, and sequences compose the
effects inferred for their subcommands. Thus, the effect of a sequence
of commands $c_1;c_2$ is the union of effects $\F_1$ and $\F_2$ of
$c_1$ and $c_2$, respectively, except that $\F_2$ is computed in a
context that includes $\F_1$. The inference rule for \C{FOREACH} takes
advantage of the $\SL$'s bind operator to lift the effect inferred for
the loop body to the level of the loop. Since records added to $\stl$
in each iteration of \C{FOREACH} are independent of the previous
iteration (recall that we make a local context independence assumption
about database programs; Sec. ~\ref{sec:opsem}), sequential
composition of the effects of different iterations is the same as
their parallel composition. Since the loop body is executed once per
each $z\in x$, the effect of the the loop is a union of effects ($\F$)
for all $z\in x$, all applied to the same state ($\stl$ and $\stg$).
That is, $\F_{loop}(\stg) = \bigcup_{z\in
x}\F_{body}(\stg)$. From the definition of the set monad's bind
operator, $\F_{loop}(\stg) = x \bind (\lambda
z.~F_{body}(\stg))$, which mirrors the definition of the rule.
%This development mirrors ideas found in type checkers for certain
%dependent type systems~\cite{KJ14}.

>>>>

The state transformer rules, like the earlier RG rules, closely follow
the corresponding reduction rules in Fig.~\ref{fig:txnimp}, except
that their language of expression is $\SL$. For instance, while the
reduction rule for \C{UPDATE} declaratively specifies the set of
updated records, the state transformer rule uses $\SL$'s bind
operation to \emph{compute} the set. Other SQL rules do likewise. The
rules for \C{LET} binders, conditionals, and sequences compose the
effects inferred for their subcommands. Thus, the effect of a sequence
of commands $c_1;c_2$ is the union of effects $\F_1$ and $\F_2$ of
$c_1$ and $c_2$, respectively, except that $\F_2$ is computed in a
context that includes $\F_1$ (we write $\F_1 \cup \F_2$ as a shorthand
for $\lambda(\stg).~\F_1(\stg) \cup \F_2(\stg)$). The inference rule
for \C{FOREACH} takes advantage of the $\SL$'s bind operator to lift
the effect inferred for the loop body to the level of the loop. Since
records added to $\stl$ in each iteration of \C{FOREACH} are
independent of the previous iteration (recall that we make a local
context independence assumption about database programs; Sec.
~\ref{sec:opsem}), sequential composition of the effects of different
iterations is the same as their parallel composition. Since the loop
body is executed once per each $z\in x$, the effect of the the loop is
a union of effects ($\F$) for all $z\in x$, all applied to the same
state ($\stl$ and $\stg$).  That is, $\F_{loop}(\stg) = \bigcup_{z\in
x}\F_{body}(\stg)$. From the definition of the set monad's bind
operator, $\F_{loop}(\stg) = x \bind (\lambda z.~F_{body}(\stg))$,
which mirrors the definition of the rule.
%This development mirrors ideas found in type checkers for certain
%dependent type systems~\cite{KJ14}.
